Manuscript_Github_Pchabaudi_compBerghei

### 2022-08-06 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_compBerghei.Rmd
Analyse berghei subfam transcription as well as chab

---
title: "Manuscript_Github_Pchabaudi_boilerplate"
author: "Timothy Little"
date: "29/07/2022"
output: html_document
---

Looking at GO enrichment of genes in- and out-of-phase between the 24 asexual life cycles of rMT and SBP transmitted P. chabaudi.

@@

cir_info <- read_delim('../standard_files/Pchabaudi_pir_info.txt',
           delim = '\t')
cir_id <- cir_info$Geneid

@@

clade_col <- setNames(
  c('#F0E442',
    '#56B4E9', #L1
    '#009E73',
    '#0072B2',
    '#E69F00', #S1
    '#D55E00',
    '#CC79A7',
    '#999999'),
  sort(
    unique(cir_info$subfam)
  )
)

@@

>>>>

---
title: "Manuscript_Github_Pchabaudi_compBerghei"
author: "Timothy Little"
date: "29/07/2022"
output: html_document
---

Comparing the P. chabaudi and P. berghei pir expression patterns for any similarities.

@@

cir_info <- read_delim('../standard_files/Pchabaudi_pir_info.txt',
           delim = '\t') %>% 
  group_by(subfam) %>% 
  mutate(subfam_total_members = n_distinct(Geneid)) %>% 
  mutate(
    subfam_num = paste0(
      subfam, 
      ' (n = ',
      subfam_total_members,
      ')'
    )
  )
cir_id <- cir_info$Geneid

@@

clade_col <- setNames(
  c('#F0E442',
    '#56B4E9', #L1
    '#009E73',
    '#0072B2',
    '#E69F00', #S1
    '#D55E00',
    '#CC79A7',
    'black'),
  sort(
    unique(cir_info$subfam)
  )
)

@@

```{r berghei_info}
bir_clade <- read_delim('../standard_files/Pberghei_pir_clades.txt', 
                        delim = '\t') %>% 
  group_by(clade) %>% 
  mutate(subfam_total_members = n_distinct(Geneid),
         subfam_num = paste0(clade, ' (n = ', n_distinct(Geneid), ')'))
bir_id <- bir_clade$Geneid

bir_clade_col <- setNames(
  c('#F0E442',
    '#56B4E9',
    '#009E73',
    '#0072B2',
    '#E69F00',
    '#D55E00',
    '#CC79A7',
    '#999999',
    'black',
    'red'),
  sort(
   unique(bir_clade$clade)
  )
)

bir_clade_col_num <- bir_clade_col
names(bir_clade_col_num) <- sapply(names(bir_clade_col), 
                               function(clade_name) {
                                 paste0(
                                   clade_name, 
                                   ' (n = ',
                                   sum(bir_clade$clade == clade_name),
                                   ')'
                                 )
                               })
```

```{r berg_tpm}
berghei_tpm <- read.csv('../chapter3_results_bergheiPir/berghei_tpm.csv', 
                        row.names = 1)
```

```{r chab_subfam_normByGeneNum}
chab_tpm_subfam.df.long <- chab_tpm.df %>% 
  rownames_to_column('Geneid') %>% 
  filter(Geneid %in% cir_id) %>% 
  melt(id.vars = 'Geneid', variable.name = 'sample', value.name = 'tpm') %>% 
  mutate(subfam = cir_info$subfam_num[match(Geneid, cir_info$Geneid)],
         transmission = str_extract(sample, pattern = 'rMT|SBP'),
         time = str_extract(sample, '[[:digit:]]{2}h|sch.culture'),
         name = paste0(time, str_extract(sample, 
                                pattern = '(_[[:digit:]]$)')))

ggplot(chab_tpm_subfam.df.long, 
       aes(x = factor(name,
                      levels = ordering(name, 
                                        order_vector = c(times_stages_table$time, 'sch.culture'))),
           y = tpm,
           fill = subfam)) +
  geom_col(position = "fill") +
  scale_fill_manual(values = clade_col_num) +
  facet_wrap(~transmission, scales = "free_x")+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90))
```

Dividing the TPM by the number of genes in each sub-family to see TPM/gene.

```{r normalised_subfam}

subfam_number.table <- group_by(cir_info, subfam_num) %>% 
  summarise(subfam_total_members = n_distinct(Geneid))

#Don't use n_distinct(Geneid) to nornalise by number of genes in a subfam group
# There has been some filtering so not all members are in the dataset.
chab_tpm_subfam_norm.df.long <- chab_tpm_subfam.df.long %>% 
  group_by(transmission, time, subfam) %>% 
         mutate(norm_tpm = tpm / cir_info$subfam_total_members[match(Geneid, cir_info$Geneid)])

ggplot(filter(chab_tpm_subfam_norm.df.long, 
              !subfam == 'ancestral (n = 1)'), 
       aes(x = factor(name,
                      levels = ordering(name, 
                                        order_vector = c(times_stages_table$time, 'sch.culture'))),
           y = norm_tpm,
           fill = subfam)) +
  geom_col(position = "stack") +
  facet_wrap(~transmission, scales = "free_x") +
  scale_fill_manual(values = clade_col_num[!names(clade_col_num) == 'ancestral (n = 1)']) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90))

```{r berghei_subfam}
berg_tpm_subfam.df.long <- berghei_tpm[,
                                       c('Geneid',
                                         str_subset(colnames(berghei_tpm), 
                                                    pattern = 'Asex'))] %>% 
  filter(Geneid %in% bir_id) %>% 
  melt(id.vars = 'Geneid', variable.name = 'sample', value.name = 'tpm') %>% 
  mutate(subfam = bir_clade$subfam_num[match(Geneid, bir_clade$Geneid)],
         experiment = str_extract(sample, pattern = '^E[[:digit:]]{2}'),
         stage = str_extract(sample, 'Asex.+(?=_[[:digit:]])'),
         name = str_extract(sample, 'Asex.+'))

ggplot(berg_tpm_subfam.df.long, 
       aes(x = factor(name,
                      levels = ordering(name,
                                        order_vector = c('Ring','Troph','Schiz','Mixed'))),
           y = tpm,
           fill = subfam)) +
  geom_col(position = "fill") +
  scale_fill_manual(values = bir_clade_col_num) +
  facet_wrap(~experiment, scales = "free_x")+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90))
```

```{r berghei_subfam_normByGeneNum}

#Don't use n_distinct(Geneid) to nornalise by number of genes in a subfam group
# There has been some filtering so not all members are in the dataset.
berg_tpm_subfam_norm.df.long <- berg_tpm_subfam.df.long %>% 
  group_by(experiment, stage, subfam) %>% 
  mutate(norm_tpm = tpm / bir_clade$subfam_total_members[match(Geneid, bir_clade$Geneid)])

ggplot(filter(berg_tpm_subfam_norm.df.long, 
              !subfam == 'ancestral (n = 1)'), 
       aes(x = factor(name,
                      levels = ordering(name, 
                                        order_vector = c('Ring','Troph','Schiz','Mixed'))),
           y = norm_tpm,
           fill = subfam)) +
  geom_col(position = "stack") +
  facet_wrap(~experiment, scales = "free_x") +
  scale_fill_manual(values = bir_clade_col_num[!names(bir_clade_col_num) %in% c('ancestral (n = 1)',
                                                                                'S6 (n = 1)')]) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90))
```
